---
title: "Hierarchical organisation simulation"
author: "Dr Tom Moore - Statistical Scientist"
date:  '`r Sys.Date()`'
knit: (function(inputFile, encoding) { out_dir <- "Reports";
      rmarkdown::render(inputFile, encoding=encoding, output_file=file.path(dirname(inputFile),
      out_dir, 'Analysis_Report_simulated.html'))})
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 5
    theme: united
    highlight: tango
    fig_width: 9
    fig_height: 6
---
<details><summary>Setup</summary>
```{r, setup, message=FALSE, warning = FALSE}
# Install packages and functions
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
    cat("Not Installed:", paste(new.pkg, collapse = ", "), "\n")
  } else {
    cat("All Packages Installed :) \n")
  }
  invisible(sapply(pkg, require, character.only = TRUE))
}

packages <- c("tidyverse","here", "simDAG","brms","cmdstanr","dagitty","ggdag","ggdist","ggtext","scales")

bayesplot::color_scheme_set("blue")                                     # Set color scheme
options(mc.cores = parallel::detectCores())                             # Parallel Chains
cmdstanr::set_cmdstan_path(path = "C:/Program Files/R/cmdstan-2.36.0")  # R-Stan
ipak(packages)                                                          # Install packages
```
</details>

# DAG
```{r, echo = FALSE, fit.width = 16, fig.height= 4, out.width='100%'}
# Create super basic DAG
simple_dag <- dagify(
  S ~ G + "T" + R + P + L + U,
  P ~ G + L + R,
  R ~ G + L,
  L ~ G + "T",
  
  G ~ U,
  R ~ U,
  L ~ U,
  
  exposure = "G",
  outcome = "S",
  latent = "U",
  coords = list(x = c(G = 1.6, "T" = 1.8, R = 2.1, S = 3, P = 2.6, L = 2.00, U = 1.7),
                y = c(G = 2.0, "T" = 1.8, R = 2.35, S = 2, P = 1.7,  L = 2.10, U = 2.7))
)

dag_plot <- simple_dag |> 
  tidy_dagitty() |> 
  mutate(var_type = case_when(
    name == "G" ~ "Exposure",
    name == "S" ~ "Outcome",
    str_detect(name, "U") ~ "Unmeasured",
    str_detect(name, "T") ~ "var",
    str_detect(name, "R") ~ "var",
    str_detect(name, "L") ~ "var",
    str_detect(name, "P") ~ "var"
  ))

ggplot(dag_plot, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges(edge_width = 0.7) +
  geom_dag_point(aes(color = var_type), alpha = 1, size = 10) +
  geom_richtext(
    aes(label = name), fontface = "bold", color = "white", size = 4,
    fill = NA, label.color = NA,
    label.padding = grid::unit(c(3, 0, 0, 0), "pt")
   ) + 
  scale_color_manual(values = c("#882255","#6699CC","#44AA99", "grey50"), guide = "none") +
  theme_dag()

ggsave(here("Analysis/Output/Figures/Figure_1.png"), width = 8, height = 4, dpi = 600)
```
Figure 1: Directed acyclic graph of the hypothesised causal relationships related to salary: G = gender; S = salary; R = role type; L = level; T = time at the organisation; and P = percentage of FTE employed. The relationship of interest is between the exposure (red) and outcome (blue). The latent variable (green) is unmeasured. Arrows indicate the direction of hypothesised causal relationships.

# Sim DAG
```{r, warning = FALSE}
# Simulate data from DAG
dag <- empty_dag() +
  
  node("u", type = "rnorm", mean = 0, sd = 1) +                              # unmeasured confounder
  node("years_since_start_date", type = "rnbinom", size = 10, prob = 0.5) +  # years since start date: a negative binomial distribution is used to ensure values are integers and positive
  node("percentage", type="rbeta", shape1 = 5, shape2 = 0.5) +               # percentage FTE employed: a beta distribution is used to ensure values are between 0 and 1
  
  node("gender", 
       type = "binomial", # A binomial distribution is used for gender (2-levels) for this minimal reproducible example.                                    
       formula = ~ u,
       betas = 0.5,       # Effect of unmeasured factor (positive = men advantaged)
       intercept = 0) +   # Men:woman ratio (0 is 50:50 - positive increases men in organisation)  
  
  node("level",
       type = "negative_binomial", # A negative binomial distribution is used for level to ensure values are integers and positive.
       formula = ~ gender + years_since_start_date + u,
       betas = c(0.005,  # Effect of level (i.e., if positive men are advantaged)
                 0.01,   # Effect of years since start date
                 0.01),  # Effect of unmeasured factor
       theta = 3,
       intercept = 0) +
  
    node("role_type", 
       type = "binomial", # Role type is either one of 2-levels - support or research                                     
       formula = ~ gender + level + u,
       betas = c(0.1,      # Effect of gender (i.e, if positive men are advantaged in support roles) # Interaction
                 0.1,      # Effect of level  (i.e, if positive then higher levels more likely to be in support roles)
                 0.5),     # Effect of unmeasured factor on level
       intercept = -0.9) + # research:support ratio (0 is 50:50 - negative indicates less people in support)

  
  node("salary", 
       type = "gaussian", # Salary is continuous and normally distributed (although in practice this will probably be skewed)
       formula = ~ gender + role_type + level + sqrt(years_since_start_date) + percentage + u, # Note the square root of years since start date is used to produce a non-linear relationship.
       betas = c(4000,2000,5000,3000,2000,5000), # Betas are listed in order: e.g., $4000 is the organisational pay gap; $2000 is the difference between support and research roles.
       intercept = 80000, 
       error = 2000)

# Set seed and simulate from DAG
set.seed(42)
sim_dat <-sim_from_dag(dag=dag, n_sim = 500)

# Add interaction between gender and role type
sim_dat <- sim_dat %>%
  mutate(salary = case_when(
    gender == "TRUE" & role_type == "TRUE"  ~ salary + 2000,
    TRUE ~ salary))

# Rename and tidy variables
sim_dat <- sim_dat %>%
  mutate(gender = ifelse(gender == "TRUE","man","woman")) %>%
  mutate(role_type = ifelse(role_type == "TRUE","support","research")) %>%
  mutate(percentage = round(percentage,2)) %>%
  mutate(level = level + 1)

# Reassign levels of gender
sim_dat$gender <- factor(sim_dat$gender, levels = c("woman", "man"))
```

# Intervention 1
```{r}
# sim_dat %>% group_by(gender) %>% summarise(median = median(salary))
# 101950.66	- 94763.87	= 7186.79

# sim_dat <- sim_dat %>%
#   mutate(salary = ifelse(gender == "woman",
#                          salary + 7187, salary))
```

# Salary error measurement calculation
```{r}
# Define salary bands
s <- c(50000, 60000, 70000, 80000, 90000, 100000, 120000, 140000, 160000, 200000)

# Create labels for salary band bins
labels <- c("below 50000", paste(head(s, -1), tail(s, -1), sep = "-"), "above 200000")

# Function to find min and max of salary band
find_bracket <- function(salary, breaks) {
  bin <- cut(salary, breaks = c(-Inf, breaks, Inf), labels = labels, right = FALSE)
  if (is.na(bin)) {
    return(c(NA, NA))
  } else {
    bounds <- strsplit(as.character(bin), "-")[[1]]
    return(as.numeric(bounds))
  }
}
```

# Intervention 2
```{r}

sim_dat <- sim_dat %>%
 rowwise() %>%
 mutate(
   min = find_bracket(salary, s)[1],
   max = find_bracket(salary, s)[2]) %>%
  
# Promote all women in research in level 1 to the next band 
  
# mutate(
#   min = ifelse(gender == "woman" & role_type == "research" & level == 1,
#                s[which(s > min)[1]],
#                min),
#   max = ifelse(gender == "woman" & role_type == "research" & level == 1, 
#                s[which(s > min)[1] + 1],
#                max)) %>%
  
 ungroup()
```

# Simulated data
```{r}
# Calculate salary estimate and standard deviation for error measurement model
sim_dat <- sim_dat %>%
  mutate(mid_est = (max + min) / 2) %>% # Calculate range mean
  mutate(sd_val  = (max - min) / 4) %>% # Calculate approximate standard deviation
  mutate(percentage_std = scale(percentage)) %>% # Scale percentage to aid model fitting
  mutate(s_lab   = paste0(gender, role_type)) %>% # Define labels for spline
  dplyr::select(c("u","gender","role_type","level","years_since_start_date","percentage_std","salary","min","max","mid_est","sd_val","s_lab")) # Select variables
  
head(sim_dat)
```

# Plot simulated data
```{r, echo = FALSE, warning = FALSE}
sim_dat %>%
  mutate(level = as.factor(level)) %>%
  mutate(gender = fct_recode(gender,"men" = "man", "women" = "woman")) %>%
  mutate(lab = paste(gender, "in", role_type)) %>%
   ggplot(aes(x = salary, fill = level)) + 
    theme_classic() +
     scale_fill_manual(values = c("#000000","#E69F00","#56B4E9","#009E73","#F0E442","#0072B2","#D55E00","#CC79A7","#999999")) +
     scale_y_continuous(limits = c(0,50), breaks = seq(0,50,10)) +
     scale_x_continuous(breaks = seq(70000, 150000, 20000), 
                     labels = function(x) paste0("$", format(x, big.mark = ","))) +
  geom_vline(xintercept = c(70000,80000,90000,100000,120000,140000), linetype = "dashed", color = "grey") +
  geom_histogram(binwidth = 5000, alpha = 0.8, color = "grey80", position = "stack") +
   annotate("segment", x = 71000 , xend = 79000, y = 50, yend = 50, color = "grey", size = 0.1, 
           arrow = arrow(length = unit(0.15, "cm"), ends = "both", type = "closed")) +
   annotate("segment", x = 81000 , xend = 89000, y = 50, yend = 50, color = "grey", size = 0.1, 
           arrow = arrow(length = unit(0.15, "cm"), ends = "both", type = "closed")) +
   annotate("segment", x = 91000 , xend = 99000, y = 50, yend = 50, color = "grey", size = 0.1, 
           arrow = arrow(length = unit(0.15, "cm"), ends = "both", type = "closed")) +
   annotate("segment", x = 101000 , xend = 119000, y = 50, yend = 50, color = "grey", size = 0.1, 
           arrow = arrow(length = unit(0.15, "cm"), ends = "both", type = "closed")) +
   annotate("segment", x = 121000 , xend = 139000, y = 50, yend = 50, color = "grey", size = 0.1, 
           arrow = arrow(length = unit(0.15, "cm"), ends = "both", type = "closed")) +  
  guides(fill = guide_legend(title = "Level", reverse = TRUE)) +
  ylab("Count") + xlab("Salary") +
  facet_wrap(~lab)

ggsave(here("Analysis/Output/Figures/Figure_2.png"), width = 8, height = 4, dpi = 600)
```

# Error measurement model
```{r}
# Model formula
mod_formula <- bf(
   mid_est | se(sd_val, sigma = TRUE) ~ 0 + Intercept + gender * role_type + mo(level) + percentage_std + # u +
    s(years_since_start_date, by = s_lab) # Spline as non-linear
  )

# Priors
priors <- c(
  prior(normal(100000, 50000), class = b, coef = Intercept),
  prior(normal(0, 10000), class = b)
)

mod <- brm(
  mod_formula,  
  family = gaussian(link = "identity"),
  prior = priors,
  data = sim_dat,  
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", 
  silent = 0,
  refresh = 2000,
  chains = 4, 
  cores = 4, 
  warmup = 1000, 
  iter = 2000,
  file = paste0(here("Analysis/Output/Models/Sims/A4.mod.error.measurement.s42.n500"))
)

# The 0 + Intercept syntax is used since variables are not centered or scaled - this ensures priors are implemented correctly.
# u is the confounding variable: if included the estimates are not confounded since the model has information on how the confounding occurs
```

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>
